detect patches

### 2024-01-25 scripts/classify_patches.Rmd
watershed bounds

```{r}
p <- '../results/suitability_maps'
r <- rast(file.path(p, 'Achnatherum_aridum1k-2024-01-10_09:16:55.tif'))
msk <- ifel(r < 0.80, NA, r)
m <- mask(r, msk)
m1 <- aggregate(m, 2, fun = 'mean')

@@

```{r}
system.time({ob <- terra::patches(m1,  directions = 4)})
system.time({z =  zonal(cellSize(ob, unit="ha"), ob, sum, as.raster=TRUE) })
z = ifel(z < 2.0235, NA, z)

plot(z)
```

@@

# nearest neighbor distance (enn_mn) - (Class) # note other method of aggregation can be used too!
## (modification) log(enn_mn), sqrt(enn_mn)
### Expectation: the number of colonized predicted patches increases with increasing distance.
#### why: this species is adept at long distance dispersal
##### test: glm(colonization(0|1) ~ enn_mn,  family = binomial)
```

>>>>

```{r}
p <- '../results/suitability_maps'
r <- rast(file.path(p, 'Acamptopappus_sphaerocephalus1k-2024-01-10_08:57:46.tif'))
msk <- ifel(r < 0.80, NA, r)
m <- mask(r, msk)
m1 <- aggregate(m, 2, fun = 'mean')

@@

We need to identify and names patches.

```{r}
system.time({ob <- terra::patches(m1,  directions = 4, allowgaps = FALSE)})
system.time({sizes =  zonal(cellSize(ob, unit="ha"), ob, sum, as.raster=TRUE) }) # use this has a mask
size_mask = ifel(sizes < 2.0235, NA, sizes) # remove areas < 5 acres.
ob <- mask(ob, z)
plot(ob)
```

```{r}

p <- '../results/suitability_maps'
f <- file.path(p, list.files(p, pattern = '1k-'))

patcheR <- functions(x){

  # read in file
  r <- rast(f)
  taxon <- gsub('1k.*$', '', basename(f))
  pout <- '../results/patches'

  # mask raster values < 0.8
  r <- mask(r, ifel(r < 0.80, NA, r))
  r1 <- aggregate(r, 2, fun = 'mean')

  # burn away rivers, and burn patches to hydrologic units ~ populations
  r1 <- mask(r1, hydro, inverse = TRUE)

  # detect patches
  pat <- terra::patches(r1,  directions = 4, allowgaps = FALSE)

  # remove patches < 5 acres
  sizes <- zonal(cellSize(pat, unit="ha"), pat, sum, as.raster=TRUE)
  size_mask <- ifel(sizes < 2.0235, NA, sizes)
  r1 <- mask(r1, size_mask)

  # resample the patches to the original raster resolution
  resample(r1, r, threads = 16, method = 'near',
           filename = file.path(pout, paste0(taxon, '.tif')))

}

@@

# nearest neighbor distance (enn_mn) - (Class) # note other method of aggregation can be used too!
## (modification) log(enn_mn), sqrt(enn_mn)
### Expectation: the number of colonized predicted patches increases with increasing distance.
#### why: this species is adept at long distance dispersal
##### test: glm(colonization(0|1) ~ enn_mn,  family = binomial)
